---
title: 'HW 3.6: Exploration'
author: "Isaac Kleisle-Murphy"
date: "3/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(psych)
source("constants.R")
source("utils.R")
source("ingest.R")

covid_df = read.csv("./data/training_data_processed.csv") %>% .[, 3:ncol(.)]
```

## 1.) Correlation plots/pairs

```{r}

feat_names = c("response", CONT_COLNAMES)
cors_with_response = cor(covid_df[, feat_names])[1, ] %>% as.numeric()
corr_df = data.frame(feature = feat_names, 
                     r = cors_with_response) %>%
  arrange(desc(cors_with_response))
  
corr_df
```
Note that a lot of the similar features are similarly correlated. Suggests PCA/Lasso/Ridge could be useful to knock down redundancy?

#### Correlation plots part 1
```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(covid_df[, corr_df %>% pull(feature) %>% .[c(1, 2:6)] %>% as.character()], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

#### Correlation plots part 2
```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(covid_df[, corr_df %>% pull(feature) %>% .[c(1, 7:11)] %>% as.character()], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

#### Correlation plots part 2
```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(covid_df[, corr_df %>% pull(feature) %>% .[c(1, 12:16)] %>% as.character()], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

#### Correlation plots part 2
```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(covid_df[, corr_df %>% pull(feature) %>% .[c(1, 17:21)] %>% as.character()], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

#### Correlation plots part 2
```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(covid_df[, corr_df %>% pull(feature) %>% .[c(1, 22:26)] %>% as.character()], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```


#### Correlation plots part 2
```{r}
# http://www.sthda.com/english/wiki/scatter-plot-matrices-r-base-graphs
pairs.panels(covid_df[, corr_df %>% pull(feature) %>% .[c(1, 26:length(.))] %>% as.character()], 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

My main takeaway here is that a lot of the facebook features are both correlated, but also skewed. Might be worth a log transform, we'll revisit in a second. 


## 2.) The Response Variable
```{r}
ggplot(covid_df, aes(x = response)) +
  geom_density()

ggplot(covid_df, aes(x = response, color=as.factor(county))) +
  geom_density()

ggplot(covid_df, aes(x = log(response))) +
  geom_density()

ggplot(covid_df, aes(x = log(response), color=as.factor(county))) +
  geom_density()
```

Would really like to have N for a binomial model here. Obviously county matters


## 3.) Seasonalities

#### Overall Plot
```{r}
covid_df = covid_df %>%
  left_join(., data.frame(date = unique(covid_df$date), date_idx = 1:length(unique(covid_df$date))),
            by=c("date"))

ggplot(covid_df, aes(x = date_idx, y = response)) + 
  geom_point() + 
  stat_smooth()

ggplot(covid_df, aes(x = date_idx, y = log(response + 1))) + 
  geom_point() + 
  stat_smooth()
```


Things have clearly changed over time (no duh); we'll need to model either with something smooth or as a varying-coefficient model. 

```{r}
ggplot(covid_df, aes(x = date_idx, y = response, color = as.factor(county))) + 
  geom_point() + 
  stat_smooth()

ggplot(covid_df, aes(x = date_idx, y = log(response), color = as.factor(county))) + 
  geom_point() + 
  stat_smooth()
```

Takeaway here is that time interacts with county: we'll probably want something like ` ~ time + county + time*county`. 


## Back to Correlations: Logging Facebook
```{r}
cors_log_response = cor(covid_df[, as.character(corr_df$feature)] %>% mutate_at(., 2:length(corr_df$feature), function(x) log(x + 1e-10)))[1, ] %>% as.numeric()

corr_log_df = data.frame(feature =  corr_df$feature, 
                         r = cors_log_response) %>%
  arrange(desc(cors_log_response))
corr_log_df
```

Nope nvm.
